---
title: "Statistical report: Eligibility for Vericiguat in realworld heart failure patients with non-worsening HFrEF"
subtitle: "Draft"
date: "`r Sys.Date()`"
cover: static/ki_logo_vert_rgb.png
editor: source
format:
  pdf:
    documentclass: scrartcl
    template-partials:
      - "static/before-body.tex"
      - "static/_titlepage.tex"
    include-in-header: 
      - "static/in-header.tex"
    toc: true
    toc-depth: 3
    toc-title: Contents
    lof: false
    lot: true
    cap-location: top
    number-sections: true
    colorlinks: false
    keep-tex: false
bibliography: static/references.bib
csl: static/bmj.csl
link-citations: true
link-bibliography: true
nocite: '@*'
knitr:
  opts_chunk: 
    dev: cairo_pdf
    collapse: true
    comment: "" 
    echo: FALSE
    include: TRUE
    warning: FALSE
    message: FALSE
    fig.pos: "H"
    fig.path: "../output/figs/"
    fig.height: 6
    fig.width: 10
    R.options:
      knitr.graphics.auto_pdf: true
---

```{r}
#| label: set-up-load-data
options(knitr.kable.NA = "")

# load packages, globals and project specific functions
source(here::here("setup/setup.R"))

# load data
load(here("data/clean-data/data.RData"))

# load workbook to write tables to Excel
wb <- loadWorkbook(here("output/tabs/tables.xlsx"))
sheets <- names(wb)
```           

\newpage

# Data

## Data sources

The Swedish Heart Failure Registry (SwedeHF) is a nationwide quality registry started in 2000, 
that includes in- and out-hospital patients with HF, regardless of EF and previous history of HF[@swedehf]. 
Around 80 variables are recorded at 
discharge from hospital or at the outpatient visit, i.e. index date, and 
entered into an electronic database managed by the Uppsala Clinical Research 
Center (Uppsala, Sweden). Up to April 2017 the only inclusion criterion was a 
clinical diagnosis of HF, which was thereafter defined according to the ICD-10 
codes I50.0, I50.1, I50.9, I42.0, I42.6, I42.7, I25.5, I11.0, I13.0, I13.2. 
Informed consent is not required, although patients are informed of entry and can opt out. 
In 2021, 69 of 76 Swedish hospitals enrolled patients in SwedeHF, which had a 32% coverage of the prevalent HF population in Sweden[@annualreport]. 

SwedeHF was linked with national Swedish registries (SHFDB v 4.1.0) through the 
Swedish personal identity number (PIN)[@pin]: 

- The National Patient Register (The National Board of Health and Welfare) for 
additional comorbidities and outcomes
- The Cause of Death Register (The National Board of Health and Welfare) for outcomes 
death
- LISA[@lisa] and the Register of the Total Population (Statistics Sweden) for 
socio-economic factors

## Acknowledgments

Please include "We thank all staff members at all care units in Sweden for their contribution to the SwedeHF register." 
in the Acknowledgements of any resulting articles. 

## Ethics

The study was approved by the Swedish Ethical Review Authority, dnr 2021-04326. 

\newpage

## Inclusion/exclusion criteria

```{r}
#| label: tbl-flow
#| tbl-cap: Inclusion/exclusion criteria
default_kable(flow) %>%
  row_spec(c(1, 10, 19), bold = T)
```

## Definitions

Information on data sources, definitions etc. are found https://kiheartfailure.github.io/shfdb4/.

{{< include src/vars.qmd >}}

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with care.

## Baseline characteristics

Baseline characteristics are presented and categorical variables are presented with n and percentage whereas 
continuous variables are presented with median [first quartile-third quartile] 

## Outcomes

The following outcomes are considered: `r paste0(outvars$name, collapse = ", ")`

Number of events, person-years and the incidence rate (96% CI) are presented. 

Data were censored at 2021-12-31, emigration from Sweden or, when it was itself not an event, death. 

QUESTION: Should cut the time at for example 5 years or want all fu time?

QUESTION: Should "First acute kidney injury", "First hyperkalemia", "First syncope", "First stroke", be hospitalizations or also out-patient care. Risk is that if NOT hospitalization will not be new (or worsening) event. Should be total (repeated) or first?

QUESTION: acute kidney injury is that : ICD:N17-9, KAS00, KAS10, KAS20, Z491, Z492 OP:DR014, DR015, DR016, DR020, DR012, DR013, DR023, DR024, TJA33, TJA35?

\clearpage

# Results

## Baseline characteristics

{{< include src/base_tab.qmd >}}

QUESTION: Hypertension is defined from ICD code wheras Hypotension is defined from sys bp. Should do in similar way?

QUESTION: Need icd codes for Hypothyroidism, Hyperthyroidism

QUESTION: How define Other CV disease?

## Outcomes

```{r}
medsurv <- rsdata %>%
  summarise(
    med = fn(median(sos_outtime_death / 365.25), dig = 1),
    min = fn(min(sos_outtime_death / 365.25), dig = 1),
    max = fn(max(sos_outtime_death / 365.25), dig = 1)
  ) %>%
  mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
  pull(out)

sumsurv <- rsdata %>%
  summarise(sumpy = fn(sum(sos_outtime_death) / 365.25, dig = 0)) %>%
  pull(sumpy)
```

The median (min-max) follow-up is 
`r medsurv` years for a total of `r sumsurv` patient-years of follow-up.

{{< include src/outcomes.qmd >}}

QUESTION: Do you want km curves or similar?

\clearpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 

https://github.com/KIHeartFailure/swedehf-vericiguat. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References

::: {#refs}
:::
